---
title: "Calcium Supplementation During Pregnancy"
author: "OD, AF"
date: "2/8/2020"
output: rmarkdown::github_document
always_allow_html: true
---

### Required packages for laoding dataset, manipulation and exploration:


```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(patchwork)
library(skimr)
library(scales)
library(knitr)
library(xtable)
library(kableExtra)
library(knitr)
library(rhdf5)
```
##### Code hidden
### Laoding dataset: [output_1.csv]


```{r echo=FALSE, message=FALSE}
draft_2 <- read_csv('Output_Files/output_1.csv',
               col_types = list(
                 item = col_character(),
                 item_outcome = col_character(),
                 gender = col_character(),
                 lbw_sg_cat = col_character(),
                 exposed_in = col_character(),
                 item_age_group = col_character(),
                 total_population_untracked = col_double(),
                 total_population_tracked = col_double(),
                 total_population = col_double(),
                 total_population_living = col_double(),
                 total_population_dead = col_double(),
                 input_draw = col_double(),
                 random_seed = col_double(),
                 simulation_run_time = col_double(),
                 calcium_supplementation_intervention.proportion = col_character(),
                 value = col_double()))
```

##### Code hidden

### Final dataset: Converting columns  [final_dataset.csv]

```{r echo=FALSE}
lbls <- c("1_to_4", "Early neonatal", "Late neonatal" , "Post neonatal")

final_dataset <- draft_2 %>% 
  mutate(gender = factor(gender, 
                         labels = c('Female', 'Male'))) %>% 
  mutate(age_group = factor(item_age_group,
                                 labels = lbls)) %>% 
  mutate(ca_supp = factor(calcium_supplementation_intervention.proportion,
                          labels = c('NO', 'YES')))
```
##### Code hidden

<br>
<br>

***

<br>
<br>

## I - Calculating Death Rate for each treatment group:


### Table 1.1: Crude death rate: CDR  [crude_death_rate.pdf]

```{r}
final_dataset %>%
  filter(item %in% c('death', 'person_time')) %>%
  group_by(input_draw, ca_supp, item ) %>% 
  summarise(sum = sum(value)) %>% 
  spread(item, sum) %>% 
  mutate(crude_death_rate = death / person_time) %>% 
  select(-c(death, person_time)) %>% 
  spread(ca_supp, crude_death_rate) %>% 
  mutate(Delta = NO - YES) %>% 
  gather(calcium_supplementation, value, 2:4) %>% 
  group_by(calcium_supplementation) %>% 
  summarise(mean_CDR = round(1000 * mean(value),1)) %>% 
  kable("html") %>% 
  kable_styling(full_width = F,font_size = 15)
```

### Table-1.2: Death rate by gender 

```{r}
final_dataset %>%
  filter(item %in% c('death', 'person_time')) %>%
  group_by(input_draw, ca_supp, item , gender) %>% 
  summarise(sum = sum(value)) %>% 
  spread(item, sum) %>% 
  mutate(death_rate = death / person_time) %>% 
  select(-c(death, person_time)) %>% 
  spread(ca_supp, death_rate) %>% 
  mutate(Delta = NO - YES) %>% 
  gather(calcium_supplementation, value, 3:5) %>% 
  group_by(gender, calcium_supplementation) %>% 
  summarise(mean_death_rate = round(1000 * mean(value),1)) %>% 
  kable("html") %>% 
  kable_styling(full_width = F,font_size = 15)
```


### Table-1.3: Death rate by age groups 

```{r}
final_dataset %>% 
  filter(item %in% c('death', 'person_time')) %>%
  group_by(input_draw, ca_supp, item , age_group) %>% 
  summarise(sum = sum(value)) %>% 
  spread(item, sum) %>% 
  mutate(death_rate = death / person_time) %>% 
  select(-c(death, person_time)) %>% 
  spread(ca_supp, death_rate) %>% 
  mutate(Delta = NO - YES) %>% 
  gather(calcium_supplementation, value, 3:5) %>% 
  group_by(age_group, calcium_supplementation) %>% 
  summarise(mean_death_rate = round(1000 * mean(value),1)) %>% 
  kable("html") %>% 
  kable_styling(full_width = F,font_size = 15)
```



### Table-1.4: Death rate by causes


```{r}
final_dataset %>% 
  filter(item %in% c('death', 'person_time')) %>%
  unite(causes, item, item_outcome) %>% 
  group_by(input_draw, ca_supp, causes) %>% 
  summarise(sum = sum(value)) %>%
  spread(causes, sum) %>% 
  rename(person_time = person_time_NA) %>%
  gather(causes, value, 3:10) %>%  
  mutate(death_rate = value / person_time) %>% 
  select(-c(value, person_time)) %>% 
  spread(ca_supp, death_rate) %>% 
  mutate(Delta = NO - YES) %>% 
  gather(calcium_supplementation, value, 3:5) %>% 
  group_by(causes, calcium_supplementation) %>% 
  summarise(mean_death_rate = round(1000 * mean(value),1)) %>% 
  kable("html") %>%
  kable_styling(full_width = F,font_size = 15)
```

<br>
<br>

***

<br>
<br>

## II - Calculating DALY for each treatment group:


#************************************************

DALY (ylls + ylds) by age groups based on calcium supplementation

Reference table from Abie

Ca_Supp |mean_DALY
--------|--------
No:     |242598.0
Yes:    |233372.0

# * * * * * * * * * * * * * * * * * * * * * * * * *


### Table-2.1: DALY (ylls + ylds) based on calcium supplementation

```{r}
final_dataset %>%
  filter(item %in% c('ylls', 'ylds', 'person_time')) %>% 
  group_by(input_draw, ca_supp, item) %>% 
  summarise(value = sum(value)) %>% 
  spread(item, value) %>% 
  mutate(total_dalys = ylds + ylls,
         dalys_per_year = total_dalys / person_time) %>% 
  select(input_draw, ca_supp, dalys_per_year) %>% 
  spread(ca_supp, dalys_per_year) %>% 
  mutate(Delta = NO - YES) %>% 
  gather(calcium_supplementation, value, 2:4) %>% 
  group_by(calcium_supplementation) %>% 
  summarise(mean_Daly = round(100000 * mean(value),0))%>% 
  kable("html") %>%
  kable_styling(full_width = F,font_size = 15)
```


### Table-2.2: Daly by gender

```{r}
final_dataset %>%
  filter(item %in% c('ylls', 'ylds', 'person_time')) %>% 
  group_by(input_draw, ca_supp, item, gender) %>% 
  summarise(value = sum(value)) %>% 
  spread(item, value) %>% 
  mutate(total_dalys = ylds + ylls,
         dalys_per_year = total_dalys / person_time) %>% 
  select(input_draw, gender, ca_supp, dalys_per_year) %>% 
  spread(ca_supp, dalys_per_year) %>% 
  mutate(Delta = NO - YES) %>% 
  gather(calcium_supplementation, value, 3:5) %>% 
  group_by(gender, calcium_supplementation) %>% 
  summarise(mean_Daly = round(100000 * mean(value),0))%>% 
  kable("html") %>%
  kable_styling(full_width = F,font_size = 15)
```


### Table-2.3: Daly by age group

```{r}
final_dataset %>%
  filter(item %in% c('ylls', 'ylds', 'person_time')) %>% 
  group_by(input_draw, ca_supp, item, age_group) %>% 
  summarise(value = sum(value)) %>% 
  spread(item, value) %>% 
  mutate(total_dalys = ylds + ylls,
         dalys_per_year = total_dalys / person_time) %>% 
  select(input_draw, age_group, ca_supp, dalys_per_year) %>% 
  spread(ca_supp, dalys_per_year) %>% 
  mutate(Delta = NO - YES) %>% 
  gather(calcium_supplementation, value, 3:5) %>% 
  group_by(age_group, calcium_supplementation) %>% 
  summarise(mean_Daly = round(100000 * mean(value),0))%>% 
  kable("html") %>%
  kable_styling(full_width = F,font_size = 15)
```


### Table-1.4: Daly by causes

```{r}
final_dataset %>% 
  filter(item %in% c('ylls', 'ylds', 'person_time')) %>%
  unite(causes, item, item_outcome) %>% 
  group_by(input_draw, ca_supp, causes) %>% 
  summarise(sum = sum(value)) %>%
  spread(causes, sum) %>% 
  rename(person_time = person_time_NA) %>% 
  gather(causes, value, 4:18) %>% 
  separate(causes, c('type', 'causes'), sep = '_', extra = 'merge') %>%
  spread(type, value) %>% 
  mutate(total_dalys = ylds + ylls,
         dalys_per_year = total_dalys / person_time) %>% 
  select(input_draw, causes, ca_supp, dalys_per_year) %>% 
  spread(ca_supp, dalys_per_year) %>% 
  mutate(Delta = NO - YES) %>% 
  gather(calcium_supplementation, value, 3:5) %>% 
  group_by(causes, calcium_supplementation) %>% 
  summarise(mean_Daly = round(100000 * mean(value),0))%>% 
  kable("html") %>%
  kable_styling(full_width = F,font_size = 15)
```

