---
title: "ca_supp_1"
author: "od"
date: "1/16/2020"
output:
  pdf_document: default
  html_document: default
---

## Loading required pacakges to laod datasets, manipulate and explore some outputs

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(patchwork)
library(skimr)
library(scales)
library(knitr)
library(xtable)
library(kableExtra)
library(knitr)
library(rhdf5)
```


## Reading datasets in R with h5ls 
```{r}
h5ls("DataSets/output.hdf")
```

## EDA 1: 
* The first file (renamed here 'axis_0') and the 'items' file are similar with dimensions = 1382
* Files label0, label1, level0 and level1 have 3998, 3998, 200 and 10 elements respectivily.
* files renamed here 'values' is the only one in array shape with dimensions (1382, 3998)

Question: how are the 4 files [label0, label1, level0 and level1] related to the other files?

```{r}
h5read("DataSets/output.hdf", name = "data/axis0")  -> axis_0
h5read("DataSets/output.hdf", name = "data/axis1_label0")  -> label0
h5read("DataSets/output.hdf", name = "data/axis1_label1")  -> label1
h5read("DataSets/output.hdf", name = "data/axis1_level0")  -> level0
h5read("DataSets/output.hdf", name = "data/axis1_level1")  -> level1
h5read("DataSets/output.hdf", name = "data/block0_items")  -> items
h5read("DataSets/output.hdf", name = "data/block0_values") -> values
```



```{r}
axis_0 %>% dim()
label0 %>% dim()
label1 %>% dim()
level0 %>% dim()
level1 %>% dim()
items %>% dim()
values %>% dim()
```

```{r}
axis_0 %>% head
label0 %>% head
label1 %>% head
level0 %>% head
level1 %>% head
items %>% head
values[1:10]
```

## Data wrangling process: 
* After viewing the file in PYTHON, I realised that 'items' file contains the rows names from the orginal dataset. 
* That is what I tried to replicate in the following section using only values file and items file.
    * first step: transpose the array (rows-to-columns)
    * second step: add 'items' containt as columns names
* output: draft_1
  

### 

```{r}
t_values <- t(values)
colnames(t_values) <- items
t_values[c(1:3), c(1:3)]
t_values[c(1:3), c(1380:1382)]
draft_1 <- as_tibble(t_values)
draft_1
```


## Saving the final dataset:
  * Getting the output_1 file took few steps:
  * step-1: grouping 9 columns (listed here) and gathering all the rest in one column 'item'.
      total_population_untracked                     
      total_population_tracked                       
      total_population                              
      total_population_living                        
      total_population_dead                          
      input_draw                                     
      random_seed                                    
      simulation_run_time                           
      calcium_supplementation_intervention.proportion
  * step-2: 
      - seperating gender, age_group ('item_age_group) where relevent
      - separating item and outcome/risk into: [item, item_outcome]
      - separate cat## and exposure year form low_birth_weight_and_short_gestation
        question: is it that useful for analysis?
        


```{r message=FALSE, warning=FALSE}
output_1 <- draft_1 %>% 
  select(contains("years_"), everything()) %>% 
  select(contains("_event_"), everything()) %>% 
  select(contains("_cases_at_sim_"), everything()) %>% 
  select(contains("_exposed_"), everything()) %>% 
  select(contains("person_time"), everything()) %>% 
  select(contains("ylls_due"), everything()) %>% 
  select(contains("ylds_due"), everything()) %>% 
  select(contains("death_due"), everything()) %>% 
  gather("item", "value", 1:1373) %>% 
  select(10, everything()) %>% 
  mutate(item = str_replace_all(item, '20_among_', '20_sep_')) %>% 
  mutate(item = str_replace_all(item, '21_among_', '21_sep_')) %>% 
  mutate(item = str_replace_all(item, '22_among_', '22_sep_')) %>% 
  mutate(item = str_replace_all(item, '23_among_', '23_sep_')) %>% 
  mutate(item = str_replace_all(item, '24_among_', '24_sep_')) %>% 
  mutate(item = str_replace_all(item, '_in_age_group_', '_sep_')) %>%
  separate(item, 
           c('item', 'item_age_group'), 
           sep = '_sep_') %>% 
  mutate(item = str_replace_all(item, '_exposed_in_20', '_sep2_20'))%>% 
  separate(item, 
           c('item', 'exposed_in'), 
           sep = '_sep2_') %>%  
  mutate(item = str_replace_all(item, 'gestation_cat', 'gestation_sep3_cat'))   %>% 
  separate(item, 
           c('item', 'lbw_sg_cat'), 
           sep = '_sep3_') %>%
  mutate(item = str_replace_all(item, '_among_male', '_sep4_male')) %>% 
  mutate(item = str_replace_all(item, '_among_female', '_sep4_female')) %>% 
  separate(item, 
           c('item', 'gender'), 
           sep = '_sep4_') %>% 
  mutate(item = str_replace_all(item, 'th_due_to_', 'th_sep5_')) %>% 
  mutate(item = str_replace_all(item, 'ds_due_to_', 'ds_sep5_')) %>% 
  mutate(item = str_replace_all(item, 'ls_due_to_', 'ls_sep5_')) %>% 
  separate(item, 
           c('item', 'item_outcome'), 
           sep = '_sep5_') 
  
```

Not all elements in 'item' column can be separateed in gender / age group.  For somes elements, there no gender or age group, which explains all the NA in those new columns.

What I was not expecting is to have NAs in the 'value' column. They are all in one item type 
(See the tbl_df ot next section)

```{r}
output_1 %>% head
output_1 %>% select(value) %>% skim
output_1 %>% filter(is.na(value)) 
```


NAs: Almost half of 'low_birth_weight_and_short_gestation' in item column have corresponding NAs in 
      value column.

```{r}
output_1%>% 
  group_by(item) %>% 
  summarise(pct = mean(is.na(value)) * 100) %>% 
  arrange(desc(pct)) 
```

## Saving output_1 files:
 * you can save the file by replacing 'Output_Files/' with the current directory.
 * can ignore this part [saving output_1.csv] for space/time sake and just run the rest of the code

```{r}
write_csv(output_1, 'Output_Files/output_1.csv')
```

```{r}
draft_2 <- read_csv('Output_Files/output_1.csv',
               col_types = list(
                 item = col_character(),
                 item_outcome = col_character(),
                 gender = col_character(),
                 lbw_sg_cat = col_character(),
                 exposed_in = col_character(),
                 item_age_group = col_character(),
                 total_population_untracked = col_double(),
                 total_population_tracked = col_double(),
                 total_population = col_double(),
                 total_population_living = col_double(),
                 total_population_dead = col_double(),
                 input_draw = col_double(),
                 random_seed = col_double(),
                 simulation_run_time = col_double(),
                 calcium_supplementation_intervention.proportion = col_character(),
                 value = col_double()))
```



Converting to factor: gender, age group, calcium supplementation and reproducing the ealier graph

```{r}
#draft_2 %>% 
#  distinct(item_age_group)

lbls <- c("1_to_4", "Early neonatal", "Late neonatal" , "Post neonatal")

final_dataset <- draft_2 %>% 
  mutate(gender = factor(gender, 
                         labels = c('Female', 'Male'))) %>% 
  mutate(age_group = factor(item_age_group,
                                 labels = lbls)) %>% 
  mutate(ca_supp = factor(calcium_supplementation_intervention.proportion,
                          labels = c('NO', 'YES')))
final_dataset %>% 
  distinct(age_group)

```

```{r}
final_dataset %>% 
  count(age_group) %>% kable()
```


## Saving final_dataset:
 * you can save the file by replacing 'Output_Files/' with the current directory.
 * can ignore this part [saving final_dataset.csv] for space/time sake and just run the rest of the code

```{r}
write_csv(final_dataset, 'Output_Files/final_dataset.csv')
```


## Example of plot  #############
```{r fig.width=8, fig.align='center'}
final_dataset %>% 
  filter(item =='death') %>% 
  filter(item_outcome == 'neonatal_preterm_birth') %>% 
  filter(item_age_group %in% c('early_neonatal',
                               'late_neonatal')) %>%
  ggplot(aes(value,
             color = ca_supp,
             fill = ca_supp))+
  geom_density(size = 1, alpha = .3)+
  facet_grid(gender ~ age_group, scales = 'free')+
  labs(title = 'Death due to neonatal preterm birth',
       subtitle = 'By age-group and by gender\n\n',
       color = 'Calcium \nSupplementation',
       x = '', y = '')+
  guides(fill = FALSE)+
  theme_bw()+
  scale_color_manual(values = c('firebrick', 'darkgreen'))+
  theme(legend.position = c(.9,1.065),
        legend.direction = 'horizontal',
        strip.text = element_text(size = 10),
        axis.text.y = element_blank(),
        plot.title = element_text(face = 'bold', size = 15),
        panel.background = element_rect(colour = 'black'))
```
######
#### rerun draft_2 to get final dataset for the following steps


```{r}
final_dataset %>% glimpse()
```

```{r}
final_dataset %>% head
```

```{r}
final_dataset %>% count(item)
```




Table-1: Distribution of death rate based on calcium supplementation

```{r}
final_dataset %>%
  filter(item %in% c('death', 'person_time')) %>%
  group_by(input_draw, ca_supp, item ) %>% 
  summarise(sum = sum(value)) %>% 
  spread(item, sum) %>% 
  mutate(crude_death_rate = death / person_time) %>% 
  group_by(ca_supp) %>%
  summarise(cdr = round(1000* mean (crude_death_rate),1))
```

#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Calculating Death Rate for each trreatment group:


# Table 1: Crude death rate: CDR  [crude_death_rate.pdf]

```{r}
final_dataset %>%
  filter(item %in% c('death', 'person_time')) %>%
  group_by(input_draw, ca_supp, item ) %>% 
  summarise(sum = sum(value)) %>% 
  spread(item, sum) %>% 
  mutate(crude_death_rate = death / person_time) %>% 
  select(-c(death, person_time)) %>% 
  spread(ca_supp, crude_death_rate) %>% 
  mutate(Delta = NO - YES) %>% 
  gather(calcium_supplementation, value, 2:4) %>% 
  group_by(calcium_supplementation) %>% 
  summarise(mean_CDR = round(1000 * mean(value),1)) 
#save_kable('Output_Files/kbl_aa.png')
#kable("html") %>% kable_styling('striped', full_width = F, 
#                                  stripe_color = 'red', font_size = 15)
```

Table-2: Death rate by gender 

```{r}
final_dataset %>%
  filter(item %in% c('death', 'person_time')) %>%
  group_by(input_draw, ca_supp, item , gender) %>% 
  summarise(sum = sum(value)) %>% 
  spread(item, sum) %>% 
  mutate(crude_death_rate = death / person_time) %>% 
  select(-c(death, person_time)) %>% 
  spread(ca_supp, crude_death_rate) %>% 
  mutate(Delta = NO - YES) %>% 
  gather(calcium_supplementation, value, 3:5) %>% 
  group_by(gender, calcium_supplementation) %>% 
  summarise(mean_CDR = round(1000 * mean(value),1)) %>% 
  kable("html") %>% kable_styling('striped', full_width = F, 
                                  stripe_color = 'red', font_size = 15)
#save_kable('Output_Files/kbl_aa.png')
```


Table-3: Death rate by age groups 

```{r}
final_dataset %>% 
  filter(item %in% c('death', 'person_time')) %>%
  group_by(input_draw, ca_supp, item , age_group) %>% 
  summarise(sum = sum(value)) %>% 
  spread(item, sum) %>% 
  mutate(crude_death_rate = death / person_time) %>% 
  select(-c(death, person_time)) %>% 
  spread(ca_supp, crude_death_rate) %>% 
  mutate(Delta = NO - YES) %>% 
  gather(calcium_supplementation, value, 3:5) %>% 
  group_by(age_group, calcium_supplementation) %>% 
  summarise(mean_CDR = round(1000 * mean(value),1)) %>% 
  kable("html") %>% kable_styling('striped', full_width = F, 
                                  stripe_color = 'red', font_size = 15)
#save_kable('Output_Files/kbl_aa.png')
```



Table-4: Death rate by causes

```{r}
final_dataset %>% 
  filter(item %in% c('death', 'person_time')) %>%
  unite(causes, item, item_outcome) %>% 
  group_by(input_draw, ca_supp, causes) %>% 
  summarise(sum = sum(value)) %>%
  spread(causes, sum) %>% 
  rename(person_time = person_time_NA) %>%
  gather(causes, value, 3:10) %>% 
  mutate(causes = str_remove(causes, 'death_')) %>% 
  mutate(crude_death_rate = value / person_time) %>% 
  select(-c(value, person_time)) %>% 
  spread(ca_supp, crude_death_rate) %>% 
  mutate(Delta = NO - YES) %>% 
  gather(calcium_supplementation, value, 3:5) %>% 
  group_by(causes, calcium_supplementation) %>%  
  summarise(mean_CDR = round(1000 * mean(value),1)) %>% 
  kable("html") %>% kable_styling('striped', full_width = F, 
                                  stripe_color = 'red', font_size = 15)
```

```{r}
final_dataset %>%  count(item)
```

#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Calculating DALY for each trreatment group:


# ************************************************

Reference table from Abie
No:   242598.0
Yes:  233372.0

# * * * * * * * * * * * * * * * * * * * * * * * * *



Table-3: DALY (ylls + ylds) by age groups based on calcium supplementation

```{r}
final_dataset %>%
  filter(item %in% c('ylls', 'ylds', 'person_time')) %>% 
  group_by(input_draw, ca_supp, item) %>% 
  summarise(value = sum(value)) %>% 
  spread(item, value) %>% 
  mutate(total_dalys = ylds + ylls,
         dalys_per_year = total_dalys / person_time) %>% 
  select(input_draw, ca_supp, dalys_per_year) %>% 
  spread(ca_supp, dalys_per_year) %>% 
  mutate(Delta = NO - YES) %>% 
  gather(calcium_supplementation, value, 2:4) %>% 
  group_by(calcium_supplementation) %>% 
  summarise(mean_Daly = round(100000 * mean(value),0))
   
```



```{r}
final_dataset %>%
  filter(item %in% c('ylls', 'ylds', 'person_time')) %>% 
  group_by(input_draw, ca_supp, item, gender) %>% 
  summarise(value = sum(value)) %>% 
  spread(item, value) %>% 
  mutate(total_dalys = ylds + ylls,
         dalys_per_year = total_dalys / person_time)  %>% 
  select(input_draw, gender, ca_supp, dalys_per_year) %>% 
  spread(ca_supp, dalys_per_year) %>% 
  mutate(Delta = NO - YES) %>% 
  gather(calcium_supplementation, value, 3:5) %>% 
  group_by(gender, calcium_supplementation) %>% 
  summarise(mean_Daly = round(100000 * mean(value),0))
```


```{r}
final_dataset %>%
  filter(item %in% c('ylls', 'ylds', 'person_time')) %>% 
  group_by(input_draw, ca_supp, item, age_group) %>% 
  summarise(value = sum(value)) %>% 
  spread(item, value) %>% 
  mutate(total_dalys = ylds + ylls,
         dalys_per_year = total_dalys / person_time)%>% 
  select(input_draw, age_group, ca_supp, dalys_per_year)%>% 
  spread(ca_supp, dalys_per_year) %>% 
  mutate(Delta = NO - YES) %>% 
  gather(calcium_supplementation, value, 3:5) %>% 
  group_by(age_group, calcium_supplementation) %>% 
  summarise(mean_Daly = round(100000 * mean(value),0))
```


```{r}
final_dataset %>% 
  filter(item %in% c('ylls', 'ylds', 'person_time')) %>%
  unite(causes, item, item_outcome) %>% 
  group_by(input_draw, ca_supp, causes) %>% 
  summarise(sum = sum(value)) %>%
  spread(causes, sum) %>% 
  rename(person_time = person_time_NA) %>% 
  gather(causes, value, 4:18) %>% 
  separate(causes, c('type', 'causes'), sep = '_', extra = 'merge') %>% 
  spread(type, value) %>% 
  mutate(total_dalys = ylds + ylls,
         dalys_per_year = total_dalys / person_time) %>% 
  select(input_draw, causes, ca_supp, dalys_per_year) %>% 
  spread(ca_supp, dalys_per_year) %>% 
  mutate(Delta = NO - YES) %>% 
  gather(calcium_supplementation, value, 3:5) %>% 
  group_by(causes, calcium_supplementation) %>% 
  summarise(mean_Daly = round(100000 * mean(value),0))
  
```



